---
title: "Interactive graphics and animated graphics"
output: 
  html_document:
    toc: true
    toc_float: true 
    theme: darkly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  gganimate = list(
    nframes = 50
  ),
  out.width = '100%'
)
```

# Data 

```{r, comment=NA, message=FALSE, echo=FALSE}
library(tidyverse)
library(magrittr)
library(patchwork)
data(penguins, package="palmerpenguins")
penguins <- penguins %>%
    rename(
         culmen_length = bill_length_mm, 
         culmen_depth = bill_depth_mm, 
         flipper_length = flipper_length_mm, 
         body_mass = body_mass_g
         ) %>% filter(!(is.na(culmen_length))) %>% filter(!(is.na(sex)))
head(penguins)
summary(penguins)
```

# Static graphic

```{r, comment=NA, message=FALSE, warning=FALSE, echo=T, fig.width=8}
library(tidyverse)
g <- ggplot(penguins, aes(x=flipper_length, y=body_mass, col=species))+
          geom_point() + theme_bw(base_size=16) 
g
```

# Interactive graphics

- allows the use to interact with the graphical information presented on the display.

- Cross filtering

    - Zoom by selecting an area of interest

    - Hover the line to get exact information.
    
# Creating Interactive plots    
    
## Method 1: ggplotly

`ggplotly`() converts your static plots to an interactive web-based version.
    
**Example 1**    

```{r, comment=NA, message=FALSE, warning=FALSE, echo=T, fig.width=8}
library(plotly)
ggplotly(g)
```

**Example 2**

```{r, comment=NA, message=FALSE, warning=FALSE, echo=T, fig.width=8}
p <- ggplot(penguins, aes(x=body_mass, color=sex)) + 
  geom_freqpoly(stat = "density") + 
  facet_wrap(~species)
ggplotly(p)
```


**Example 3**


```{r, comment=NA}
library(gapminder)
head(gapminder, 25)
p1 <- ggplot(gapminder, aes(gdpPercap, pop, color = continent, frame = year)) +
  geom_point(aes(size = pop, ids = country)) + 
   geom_smooth(method = "lm", se=FALSE) +
  scale_x_log10() +
  theme_bw()

ggplotly(p1)
```

**Example 4**

```{r, comment=NA, message=FALSE, warning=FALSE}
p2 <- ggplot(gapminder, aes(gdpPercap, lifeExp, size = pop, colour = country, frame=year)) +
  geom_point(alpha = 0.7, show.legend = FALSE) +
  scale_colour_manual(values = country_colors) +
  scale_size(range = c(2, 12)) +
  scale_x_log10() +
  facet_wrap(~continent) 

ggplotly(p2)
```







## Method 2: plotly

 - powered by the JavaScript library plotly.js
 
 - Mapping variables: Instead of `aes` use `~`
 
 - plotly finds a suitable geometric representation for us.

- Users can also define geometries.

- Functions to encode data

    - color: map each level to a different colour
    - colors: specify a range of colours
    - stroke
    - span
    - symbol
    - linetype
 
 **Example 1 - 1**
 

```{r, comment=NA, message=FALSE, warning=FALSE, echo=T, fig.height=5}
plot_ly(penguins, x = ~species)
```

```{r, comment=NA, message=FALSE, warning=FALSE, echo=T, fig.height=5}
plot_ly(penguins, x = ~species, y= ~island)
```

```{r, comment=NA, message=FALSE, warning=FALSE, echo=T, fig.height=5}
plot_ly(
  penguins, 
  x = ~species, 
  color = "green"
) # Doesn't give green bars

plot_ly(
  penguins, 
  x = ~species, 
  color = I("green"), 
  stroke = I("black"), 
  span = I(2)
)
```

**Example 2 - 1**


```{r, comment=NA, message=FALSE, warning=FALSE, echo=T}
library(magrittr)
plot_ly(penguins, x = ~flipper_length, y = ~body_mass, color=~species)
```

**Example 2 - 2**

There are a family of `add_*` functions.

```{r, comment=NA, message=FALSE, warning=FALSE, echo=T}
library(magrittr)
plot_ly(penguins, x = ~flipper_length, y = ~body_mass) %>%  add_markers(color=~species)
```

**Example 3**

```{r, comment=NA, message=FALSE, warning=FALSE, echo=T}
library(magrittr)
plot_ly(penguins, x = ~flipper_length, y = ~body_mass, z = ~culmen_depth, color=~species)
```

```{r, comment=NA, message=FALSE, warning=FALSE, echo=T}
library(magrittr)
plot_ly(penguins, x = ~flipper_length, y = ~body_mass, z = ~culmen_depth) %>% add_markers(color=~species)
```

```{r, comment=NA, message=FALSE, warning=FALSE, echo=T}
library(magrittr)
df <- data.frame(x=1:100, y=rnorm(100))
plot_ly(df, x = ~x, y = ~y) 
```


```{r, comment=NA, message=FALSE, warning=FALSE, echo=T}
library(magrittr)
df <- data.frame(x=1:100, y=rnorm(100))
plot_ly(df, x = ~x, y = ~y)  %>%
  add_lines(color = I("black"))
```

```{r, comment=NA, message=FALSE, warning=FALSE, echo=T}
plot_ly(df, x = ~x, y = ~y)  %>%
   add_trace(y = ~y,  mode = 'markers')
```

```{r, comment=NA, message=FALSE, warning=FALSE, echo=T}
plot_ly(df, x = ~x, y = ~y)  %>%
   add_trace(y = ~y,  mode = 'line')
```

```{r, comment=NA, message=FALSE, warning=FALSE, echo=T}
plot_ly(df, x = ~x, y = ~y)  %>%
   add_trace(y = ~y,   mode = 'lines+markers')
```

# Modify graphs: add_*

> add_histogram

```{r, comment=NA, message=FALSE, warning=FALSE, echo=T}
penguins %>% plot_ly() %>% add_histogram(x=~body_mass)

```


> add_bars (requires pre calculated counts)

```{r, comment=NA, message=FALSE, warning=FALSE, echo=T}
penguins %>% dplyr::count(species) %>%
  plot_ly() %>% add_bars(x=~species, y=~n)

```


```{r, comment=NA, echo=T, comment=NA, message=FALSE, warning=FALSE, fig.height=5}
library(dplyr)

penguins %>%
  plot_ly(x = ~species) %>% 
  add_histogram() %>%
  group_by(species) %>%
  summarise(n = n()) %>%
  add_text(
    text = ~scales::comma(n), y = ~n, 
    textposition = "top middle", 
    cliponaxis = FALSE
  )
```



# A Grammar of Animated Graphics

gganimate is an extension of the ggplot2 package for creating animated ggplots.

```{r, comment=NA, warning=FALSE, message=FALSE}
library(ggplot2)
library(gganimate)

g <- ggplot(gapminder, aes(gdpPercap, lifeExp, size = pop, colour = country)) +
  geom_point(alpha = 0.7, show.legend = FALSE) +
  scale_colour_manual(values = country_colors) +
  scale_size(range = c(2, 12)) +
  scale_x_log10() +
  facet_wrap(~continent) 
g

```

```{r, comment=NA, message=FALSE, warning=FALSE, eval=FALSE}
ggplot(gapminder, aes(gdpPercap, lifeExp, size = pop, colour = country)) +
  geom_point(alpha = 0.7, show.legend = FALSE) +
  scale_colour_manual(values = country_colors) +
  scale_size(range = c(2, 12)) +
  scale_x_log10() +
  facet_wrap(~continent) +
  # Here comes the gganimate specific bits
  labs(title = 'Year: {frame_time}', x = 'GDP per capita', y = 'life expectancy') +
  transition_time(year) +
  ease_aes('linear')

```



