---
title: "ASP 460 2.0 Special Topics in Statistics"
subtitle: "Interactive graphics"
author: "Thiyanga Talagala"
date: "2020-06-26"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: 
      - default
      - default-fonts
      - duke-blue
      - hygge-duke
      - libs/cc-fonts.css
      - libs/figure-captions.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

# Data 

```{r, comment=NA, message=FALSE, echo=FALSE}
library(tidyverse)
library(magrittr)
library(patchwork)
data(penguins, package="palmerpenguins")
penguins <- penguins %>%
    rename(
         culmen_length = bill_length_mm, 
         culmen_depth = bill_depth_mm, 
         flipper_length = flipper_length_mm, 
         body_mass = body_mass_g
         ) %>% filter(!(is.na(culmen_length))) %>% filter(!(is.na(sex)))
head(penguins)
summary(penguins)
```

---

# Static graphic

```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE, fig.width=8}
library(tidyverse)
g <- ggplot(penguins, aes(x=flipper_length, y=body_mass, col=species))+
          geom_point() + theme_bw(base_size=16) 
g
```

---
# Interactive graphics

- allows the use to interact with the graphical information presented on the display.

- Cross filtering

    - Zoom by selecting an area of interest

    - Hover the line to get exact information.

---
# Interactive plot

```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE, fig.width=8}
library(plotly)
ggplotly(g)
```

---
# Interactive plot (cont.)

```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE, fig.width=8}
p <- ggplot(penguins, aes(x=body_mass, color=sex)) + 
  geom_freqpoly(stat = "density") + 
  facet_wrap(~species)
ggplotly(p)
```

---
class: duke-orange, center, middle

# plotly

 powered by the JavaScript library plotly.js
---

# Introduction to `plot_ly()`:

- Mapping variables: Instead of `aes` use `~`

```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE, fig.height=5}
plot_ly(penguins, x = ~species)
```

---

# Introduction to `plot_ly()`

```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE}
plot_ly(penguins, x = ~flipper_length, y = ~body_mass)
```


---
# Introduction to `plot_ly()`

```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE}
plot_ly(penguins, x = ~flipper_length, y = ~body_mass, col=~species)
```

---

# Introduction to `plot_ly()`

- plotly finds a suitable geometric representation for us.

- Users can also define geometries.

- Functions to encode data

    - color: map each level to a different colour
    - colors: specify a range of colours
    - stroke
    - span
    - symbol
    - linetype
---
# `color` argument


```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE}
plot_ly(penguins, x = ~flipper_length, y = ~body_mass, color=~species)
```

---

# `colors` argument

.pull-left[
```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE, fig.height=5}
plot_ly(penguins, x = ~flipper_length, y = ~body_mass,
        color=~culmen_length)
```

Default
]

.pull-right[
```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE, fig.height=5}
plot_ly(penguins, x = ~flipper_length, y = ~body_mass,
        color=~culmen_length, colors="Accent")
```

colors="Accent"
]
---

# Define a colour: I()

```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE}
plot_ly(penguins, x=~species, color=I("red"))

```

---

# layout()

> functional approach to a layered grammar of graphics

> input: plotly object

```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE, fig.height=5}
layout(
plot_ly(penguins, x=~species), 
title="Composition of species"
)

```

---
# Works with %>% operator

```r
penguins %>% plot_ly(x = ~species) %>%
layout(title="Composition of species")
```

```{r, comment=NA, message=FALSE, warning=FALSE, fig.height=5, echo=FALSE}
penguins %>% plot_ly(x = ~species) %>% layout(title="Composition of species")
```

---

# Modify graphs: add_*

> add_histogram

```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE}
penguins %>% plot_ly() %>% add_histogram(x=~body_mass)

```

---

# Modify graphs

> add_bars (requires pre calculated counts)

```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE}
penguins %>% dplyr::count(species) %>%
  plot_ly() %>% add_bars(x=~species, y=~n)

```

---
# Modify graphs (cont.)

```{r, comment=NA, echo=FALSE, comment=NA, message=FALSE, warning=FALSE, fig.height=5}
library(dplyr)

penguins %>%
  plot_ly(x = ~species) %>% 
  add_histogram() %>%
  group_by(species) %>%
  summarise(n = n()) %>%
  add_text(
    text = ~scales::comma(n), y = ~n, 
    textposition = "top middle", 
    cliponaxis = FALSE
  )
```


---

# Line graphs

```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE}
library(plotly)
data(economics, package = "ggplot2")

# sort economics by psavert, just to 
# show difference between paths and lines
p <- economics %>%
  arrange(psavert) %>%
  plot_ly(x = ~date, y = ~psavert) %>%
  add_lines()
p
```

---
# Boxplots


```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE}

p <- penguins %>%
  plot_ly(y = ~body_mass, color = I("black"), alpha = 0.1) 
```

.pull-left[

```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE}

p %>% add_boxplot(x = "Overall")
```

]

.pull-right[

```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE, fig.width=4}
 p %>% add_boxplot(x = ~species)
```


]

---
# Boxplot

```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE}
plot_ly(penguins, x = ~body_mass, y = ~interaction(species, sex)) %>%
  add_boxplot(color = ~species) %>%
  layout(yaxis = list(title = ""))

```

---

# Combining multiple views

- Static or dynamic

- Scatterplot matrices

- Higer dimensional displayes 

- Arrange multiple plots in a single plane
---
# Combining multiple views

`subplot`: merge multiple plotly objects.

```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE}

p <- penguins %>%
  plot_ly(y = ~body_mass, color = I("black"), alpha = 0.1) 
p1 <- p %>% add_boxplot(x = "Overall")
p2 <- p %>% add_boxplot(x = ~species)
subplot(
  p1, p2, shareY = TRUE,
  widths = c(0.2, 0.8), margin = 0
) %>% hide_legend()
```
---

# Combining multiple views (cont.)

- `splom`

```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE}

dims <- dplyr::select_if(penguins, is.numeric)
dims <- purrr::map2(dims, names(dims), ~list(values=.x, label=.y))
plot_ly(
  type = "splom", dimensions = setNames(dims, NULL), 
  showupperhalf = FALSE, diagonal = list(visible = FALSE)
)
```

---

# Flexdashboard

[Click here](https://thiyanga.netlify.app/post/covid19/)

[Clickhere](/slides/dashboard_penguins.html)
