---
title: "STA 479 2.0 Special Topics in Statistics"
author: "Thiyanga Talagala"
date: "2020 - 06 - 10"
output:
  xaringan::moon_reader:
    chakra: libs/remark-latest.min.js
    lib_dir: libs
    css: 
      - default
      - default-fonts
      - duke-blue
      - hygge-duke
      - libs/cc-fonts.css
      - libs/figure-captions.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

# Individual Time Series: Points

.pull-left[
```r
library(mozzie)
library(ggplot2)
data(mozzie)
colombo.dengue <- mozzie[, 1:4]
ggplot(colombo.dengue, aes(x=ID, y=Colombo))+
  geom_point()+xlab("Time")+
  ylab("Number of dengue cases")
```

]

.pull-right[
```{r, comment=NA, fig.cap='Figure 1', echo=FALSE}
library(mozzie)
library(ggplot2)
data(mozzie)
colombo.dengue <- mozzie[, 1:4]
ggplot(colombo.dengue, aes(x=ID, y=Colombo))+
  geom_point()+xlab("Time")+
  ylab("Number of dengue cases")
```

]

This is NOT a scatterplot. Why?


- the points are spaced equally along the x-axis.

- there is a order among points. 

To emphasize time dependent relationship we can connect neighboring points with lines.

---

# Individual Time Series: Points and Lines

.pull-left[
```r
library(mozzie)
library(ggplot2)
data(mozzie)
colombo.dengue <- mozzie[, 1:4]
ggplot(colombo.dengue, aes(x=ID, y=Colombo))+
  geom_point()+
*  geom_line()+
  xlab("Time")+
  ylab("Number of dengue cases")
```

]

.pull-right[
```{r, comment=NA, fig.cap='Figure 1', echo=FALSE}
library(mozzie)
library(ggplot2)
data(mozzie)
colombo.dengue <- mozzie[, 1:4]
ggplot(colombo.dengue, aes(x=ID, y=Colombo))+
  geom_point()+
  geom_line()+
  xlab("Time")+
  ylab("Number of dengue cases")
```

]

- Lines do not represent observed data. Lines are meant as a guide to the eye.

- If few observed values a far apart or unevenly spaced, it is not suitable to connect points with lines.

---
## Individual Time Series: Lines only

.pull-left[
```r
ggplot(colombo.dengue, aes(x=ID, y=Colombo))+
*  geom_line()+
  xlab("Time")+
  ylab("Number of dengue cases")
```

]

.pull-right[
```{r, comment=NA, fig.cap='Figure 1', echo=FALSE}
ggplot(colombo.dengue, aes(x=ID, y=Colombo))+
  geom_line()+
  xlab("Time")+
  ylab("Number of dengue cases")
```

]

- Without points more emphasis is given on the overall trend and less on individual values.

- In general, when there are too many points it is better to plot without points.

- Recap: Week 4 lecture How to plot time series plots easily using `autoplot` function in the `forecast` package.

---

# Individual time series: Fill the area under the curve

.pull-left[
```r
ggplot(colombo.dengue, aes(x=ID, y=Colombo))+
*  geom_area()+
  xlab("Time")+
  ylab("Number of dengue cases")
```

]

.pull-right[
```{r, comment=NA, fig.cap='Figure 1', echo=FALSE}
ggplot(colombo.dengue, aes(x=ID, y=Colombo))+
  geom_area()+
  xlab("Time")+
  ylab("Number of dengue cases")
```

]

- Visually separates the area above and below the curve.

- More emphasis is given to the overarching trend in the series. 

- This visualization is only valid if the y axis starts at zero.

---



# Smoothing


---
# Visualising multiple time series

```{r, echo=FALSE, comment=NA,message=FALSE, warning=FALSE}
library(tidyverse)
library(magrittr)
mozzie <- mozzie[,1:5]
district_count <- mozzie %>%
 pivot_longer(-c("ID", "Year", "Week"), names_to = "District", values_to = "count")
```

.pull-left[

```{r,comment=NA, echo=FALSE}
ggplot(district_count, aes(x = ID, y = count, colour = District))+geom_point()

```

Difficult to read.

]

.pull-right[

```{r,comment=NA,echo=FALSE}
ggplot(district_count, aes(x = ID, y = count, colour = District))+geom_line()

```

By connecting points with lines we help the reader to follow the paths of each individual time series.

]

---
