---
title: "ASP 460 2.0 Data Visualization"
subtitle: "Visualizing time series data"
author: "Thiyanga Talagala"
date: ""
output:
  xaringan::moon_reader:
    chakra: libs/remark-latest.min.js
    lib_dir: libs
    css: 
      - default
      - default-fonts
      - duke-blue
      - hygge-duke
      - libs/cc-fonts.css
      - libs/figure-captions.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

# Individual Time Series: Points

.pull-left[
```r
library(mozzie)
library(ggplot2)
data(mozzie)
colombo.dengue <- mozzie[, 1:4]
ggplot(colombo.dengue, aes(x=ID, y=Colombo))+
  geom_point()+xlab("Time")+
  ylab("Number of dengue cases")
```

]

.pull-right[
```{r, comment=NA, fig.cap='Figure 1', echo=FALSE}
library(mozzie)
library(ggplot2)
data(mozzie)
colombo.dengue <- mozzie[, 1:4]
ggplot(colombo.dengue, aes(x=ID, y=Colombo))+
  geom_point()+xlab("Time")+
  ylab("Number of dengue cases")
```

]

This is NOT a scatter plot. Why?


- the points are spaced equally along the x-axis.

- there is a order among points. 

To emphasize time dependent relationship we can connect neighboring points with lines.

---

# Individual Time Series: Points and Lines

.pull-left[
```r
library(mozzie)
library(ggplot2)
data(mozzie)
colombo.dengue <- mozzie[, 1:4]
ggplot(colombo.dengue, aes(x=ID, y=Colombo))+
  geom_point()+
*  geom_line()+
  xlab("Time")+
  ylab("Number of dengue cases")
```

]

.pull-right[
```{r, comment=NA, fig.cap='Figure 1', echo=FALSE}
library(mozzie)
library(ggplot2)
data(mozzie)
colombo.dengue <- mozzie[, 1:4]
ggplot(colombo.dengue, aes(x=ID, y=Colombo))+
  geom_point()+
  geom_line()+
  xlab("Time")+
  ylab("Number of dengue cases")
```

]

- Lines do not represent observed data. Lines are meant as a guide to the eye.

- If few observed values a far apart or unevenly spaced, it is not suitable to connect points with lines.

---
## Individual Time Series: Lines only

.pull-left[
```r
ggplot(colombo.dengue, aes(x=ID, y=Colombo))+
*  geom_line()+
  xlab("Time")+
  ylab("Number of dengue cases")
```

]

.pull-right[
```{r, comment=NA, fig.cap='Figure 1', echo=FALSE}
ggplot(colombo.dengue, aes(x=ID, y=Colombo))+
  geom_line()+
  xlab("Time")+
  ylab("Number of dengue cases")
```

]

- Without points more emphasis is given on the overall trend and less on individual values.

- In general, when there are too many points it is better to plot without points.

- Recap: Week 4 lecture How to plot time series plots easily using `autoplot` function in the `forecast` package.

---

# Individual time series: Fill the area under the curve

.pull-left[
```r
ggplot(colombo.dengue, aes(x=ID, y=Colombo))+
*  geom_area()+
  xlab("Time")+
  ylab("Number of dengue cases")
```

]

.pull-right[
```{r, comment=NA, fig.cap='Figure 1', echo=FALSE}
ggplot(colombo.dengue, aes(x=ID, y=Colombo))+
  geom_area()+
  xlab("Time")+
  ylab("Number of dengue cases")
```

]

- Visually separates the area above and below the curve.

- More emphasis is given to the overarching trend in the series. 

- This visualization is only valid if the y axis starts at zero.


---
# Visualising multiple time series

```{r, echo=FALSE, comment=NA,message=FALSE, warning=FALSE}
library(tidyverse)
library(magrittr)
mozzie <- mozzie[,1:5]
district_count <- mozzie %>%
 pivot_longer(-c("ID", "Year", "Week"), names_to = "District", values_to = "count")
```

.pull-left[

```{r,comment=NA, echo=FALSE}
ggplot(district_count, aes(x = ID, y = count, colour = District))+geom_point()

```

Difficult to read.

]

.pull-right[

```{r,comment=NA,echo=FALSE}
ggplot(district_count, aes(x = ID, y = count, colour = District))+geom_line()

```

By connecting points with lines we help the reader to follow the paths of each individual time series.

]

---
# Smoothing

```{r}
library(gapminder)
head(gapminder)
```

---
# Smoothing

```{r, comment=NA, message=FALSE, echo=FALSE}
library(tidyverse)
gapminder_af <- gapminder %>%
  filter(continent == "Africa")
ggplot(gapminder_af, aes(x=year, y=lifeExp))+geom_point()+geom_line()+facet_wrap(~country)
```

---
# Smoothing

```{r, comment=NA, message=FALSE, echo=FALSE}
library(tidyverse)
gapminder_af <- gapminder %>%
  filter(continent == "Africa")
ggplot(gapminder_af, aes(x=year, y=lifeExp))+geom_line()+facet_wrap(~country)+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
---
# Smoothing

```{r, comment=NA, message=FALSE, echo=FALSE}
library(tidyverse)
gapminder_af <- gapminder %>%
  filter(continent == "Africa")
ggplot(gapminder_af, aes(x=year, y=lifeExp))+geom_smooth()+facet_wrap(~country)+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

---
# Time series plots

```{r, comment=NA, message=FALSE, warning=FALSE, fig.height=5, echo=FALSE}
library(forecast)
library(fpp2)
mozcol <- ts(mozzie$Colombo, frequency = 52, start = c(2008, 52))
autoplot(mozcol) +
  ggtitle("Dengue Count - Colombo") +
  ylab("Count") +
  xlab("Year")
```

---
# Seasonal plots

```{r, comment=NA, message=FALSE, warning=FALSE, fig.height=5, echo=FALSE}
ggseasonplot(mozcol, year.labels=TRUE, year.labels.left=TRUE) +
  ylab("Count") +
  ggtitle("Seasonal plot: Dengue Count - Colombo")
```

---
# Polar seasonal plot

```{r, comment=NA, message=FALSE, warning=FALSE, fig.height=5, echo=FALSE}
ggseasonplot(mozcol, polar=TRUE) +
  ylab("Count") +
  ggtitle("Polar seasonal plot: Dengue Count - Colombo")
```

---
# Time series plot

```{r, comment=NA, message=FALSE, warning=FALSE, fig.height=5, echo=FALSE}
autoplot(a10) +
  ggtitle("Antidiabetic drug sales") +
  ylab("$ million") +
  xlab("Year")
```
---
# Seasonal plot

```{r, comment=NA, message=FALSE, warning=FALSE, fig.height=5, echo=FALSE}
ggseasonplot(a10, year.labels=TRUE, year.labels.left=TRUE) +
  ylab("$ million") +
  ggtitle("Seasonal plot: antidiabetic drug sales")
```

---

# Polar seasonal plot

```{r, comment=NA, message=FALSE, warning=FALSE, fig.height=5, echo=FALSE}
ggseasonplot(a10, polar=TRUE) +
  ylab("$ million") +
  ggtitle("Polar seasonal plot: antidiabetic drug sales")
```
---

# Seasonal subseries plots

```{r, comment=NA, message=FALSE, warning=FALSE, fig.height=5, echo=FALSE}
ggsubseriesplot(a10) +
  ylab("$ million") +
  ggtitle("Seasonal subseries plot: antidiabetic drug sales")
```


---

```{r, comment=NA}
a10
```
---

# Seasonal subseries plots

```{r, comment=NA, message=FALSE, warning=FALSE, fig.height=5, echo=FALSE}
ggsubseriesplot(mozcol) +
  ylab("Count") +
  ggtitle("Seasonal subseries plot: Dengue count")
```

---
# Lag plots: a10

Monthly anti-diabetic drug sales in Australia from 1991 to 2008.

```{r, comment=NA, message=FALSE, echo=FALSE}
gglagplot(a10)
```

---

# Lag plots: mozzie 

Dengue counts - colombo

```{r, comment=NA, message=FALSE, echo=FALSE}
gglagplot(mozcol)
```

---

# Lag plots: ausbeer

Monthly Australian beer production: Jan 1991 – Aug 1995.

```{r, comment=NA, message=FALSE, echo=FALSE}
beer2 <- window(ausbeer, start=1992)
gglagplot(beer2)
```