---
output: 
  html_document:
    toc: true
    toc_float: true 
    theme: darkly
---

# Flipped Classroom Exercises

# What is a flipped classroom?

<div class="alert alert-info">
  <strong>"A flipped classroom is an instructional strategy and a type of blended learning, which aims to increase student engagement and learning by having students complete readings at their home and work on live problem-solving during class time ." Source: </strong>https://en.wikipedia.org/wiki/Flipped_classroom
</div>

# Learning objectives

i. Create data graphics using the ggplot2 package

ii. Identify suitable aesthetics and geoms to create plots using `qplot` and `ggplot`

iii. Identify what faceting is and apply faceting in ggplot

iv. Customize data graphics using ggplot



# Learning materials

Check the video lectures in Google Classroom.



# Data set

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This question is based on `mpg` dataset in the `ggplot2` package. Load the `ggplot2` package and `mpg` dataset using the following command.

```{r, comment=NA}
library(ggplot2)
data(mpg)
```


<span style="color:green">Fuel economy data from 1999 to 2008 for 38 popular models of cars</span>.


This dataset contains a subset of the fuel economy data that the EPA makes available on https://fueleconomy.gov/. It contains only models which had a new release every year between 1999 and 2008 - this was used as a proxy for the popularity of the car. To find out more information about the dataset you can run the command

```{r, comment=NA}
help(mpg) # or ?mpg
```

You can observe the data structure using the following command

```{r, comment=NA}
str(mpg)
```

To get a summary use the command

```{r, comment=NA}
summary(mpg)
```

# Exercise

## Write R codes to reproduce the following plots using both the `qplot` function and the `ggplot` function in the `ggplot2` package. The first one is done for you.

Help:

1. Identify the `x` variable, `y` variable, `color` variable.

2. Identify the geom type: `point`, `hist`, `smooth`, `density`

3. Identify the attribute: `color`, `shape`, etc.

## Plot 1

```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE}
qplot(displ, hwy, data=mpg, geom="point")

```

### qplot command:

```{r, comment=NA, message=FALSE, warning=FALSE, eval=FALSE}
qplot(displ, hwy, data=mpg, geom="point")

```


### ggplot command:

```{r, comment=NA, message=FALSE, warning=FALSE, eval=FALSE}
## method 1
ggplot(data=mpg aes(x=displ, y=hwy)) + geom_point()
## method 2
ggplot(data=mpg) + geom_point(aes(x=displ, y=hwy))
```

## Plot 2

```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE}
qplot(displ, hwy, color=drv, data=mpg, geom="point")

```

### qplot command:

--------

---------

### ggplot command:

---------

-------

## Plot 3

```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE}
qplot(displ, hwy, data=mpg, geom=c("point", "smooth"))

```

### qplot command:

--------

---------

### ggplot command:

---------

-------

## Plot 4

```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE}
qplot(hwy, data=mpg)

```

### qplot command:

--------

---------

### ggplot command:

---------

-------

## Plot 5

```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE}
qplot(hwy, fill=drv, data=mpg)

```

### qplot command:

--------

---------

### ggplot command:

---------

-------

## Plot 6

```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE}
qplot(displ, hwy, data=mpg, facets = .~drv)

```

### qplot command:

--------

---------

### ggplot command:

---------

-------


## Plot 6

```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE}
qplot(displ, hwy, data=mpg, facets = drv ~ .)

```

### qplot command:

--------

---------

### ggplot command:

---------

-------

## Plot 7

```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE}
qplot(hwy, data=mpg, facets = drv ~ ., binwidth=2)

```

### qplot command:

--------

---------

### ggplot command:

---------

-------

## Plot 8


```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE}
qplot(hwy, color=drv, data=mpg, geom="density")

```

### qplot command:

--------

---------

### ggplot command:

---------

-------

## Plot 9




```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE}
qplot(hwy, cty, data=mpg, geom=c("point", "smooth"),
      method="lm", facets = .~drv)

```

### qplot command:

--------

---------

### ggplot command:

---------

-------

## Plot 10



```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE}

ggplot(data=mpg, aes(y=hwy, x=cty)) + geom_point() + facet_wrap(drv~., ncol=3) + labs(x="number of cylinders") + 
  labs(y="highway miles per gallon") + labs(title="Scatterplot of highway miles per gallon vs number of cylinders")
```


ggplot command only:

--------

------