---
title: "ASP 460 2.0 Data Visualization"
author: "Dr Thiyanga Talagala"
date: "Visualizing Time Series Data"
output:
  xaringan::moon_reader:
    chakra: libs/remark-latest.min.js
    lib_dir: libs
    css: 
      - default
      - default-fonts
      - duke-blue
      - hygge-duke
      - libs/cc-fonts.css
      - libs/figure-captions.css
      - xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
---
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent(
  base_color = "#081d58",
  text_bold_color = "#ce1256",
  title_slide_text_color = "#edf8b1",
  header_font_google = google_font("Josefin Sans"),
  base_font_size = "20px",
  text_font_size = "1.5rem"
 #text_font_google   = google_font("Montserrat", "300", "300i")
 # code_font_google   = google_font("Fira Mono")
)
```

# Individual Time Series: Points

.pull-left[
```r
library(mozzie)
library(ggplot2)
data(mozzie)
colombo.dengue <- mozzie[, 1:4]
ggplot(colombo.dengue, aes(x=ID, y=Colombo))+
  geom_point()+xlab("Time")+
  ylab("Number of dengue cases")
```

]

.pull-right[
```{r, comment=NA, fig.cap='Figure 1', echo=FALSE}
library(mozzie)
library(ggplot2)
data(mozzie)
colombo.dengue <- mozzie[, 1:4]
ggplot(colombo.dengue, aes(x=ID, y=Colombo))+
  geom_point()+xlab("Time")+
  ylab("Number of dengue cases")
```

]

---

# Individual Time Series: Points

.pull-left[
```r
library(mozzie)
library(ggplot2)
data(mozzie)
colombo.dengue <- mozzie[, 1:4]
ggplot(colombo.dengue, aes(x=ID, y=Colombo))+
  geom_point()+xlab("Time")+
  ylab("Number of dengue cases")
  
```

This is NOT a scatter plot. Why?

]

.pull-right[
```{r, comment=NA, fig.cap='Figure 1', echo=FALSE}
library(mozzie)
library(ggplot2)
data(mozzie)
colombo.dengue <- mozzie[, 1:4]
ggplot(colombo.dengue, aes(x=ID, y=Colombo))+
  geom_point()+xlab("Time")+
  ylab("Number of dengue cases")
```

]


---


# Individual Time Series: Points

.pull-left[

**This is NOT a scatter plot. Why?**


- the points are spaced equally along the x-axis.

- there is a order among points. 

To emphasize time dependent relationship we can connect neighboring points with lines.



]

.pull-right[
```{r, comment=NA, fig.cap='Figure 1', echo=FALSE}
library(mozzie)
library(ggplot2)
data(mozzie)
colombo.dengue <- mozzie[, 1:4]
ggplot(colombo.dengue, aes(x=ID, y=Colombo))+
  geom_point()+xlab("Time")+
  ylab("Number of dengue cases")
```

]

---

# Individual Time Series: Points and Lines

.pull-left[
```r
library(mozzie)
library(ggplot2)
data(mozzie)
colombo.dengue <- mozzie[, 1:4]
ggplot(colombo.dengue, aes(x=ID, y=Colombo))+
  geom_point()+
*  geom_line()+
  xlab("Time")+
  ylab("Number of dengue cases")
```

]

.pull-right[
```{r, comment=NA, fig.cap='Figure 1', echo=FALSE}
library(mozzie)
library(ggplot2)
data(mozzie)
colombo.dengue <- mozzie[, 1:4]
ggplot(colombo.dengue, aes(x=ID, y=Colombo))+
  geom_point()+
  geom_line()+
  xlab("Time")+
  ylab("Number of dengue cases")
```

]

---

# Individual Time Series: Points and Lines

.pull-left[


- Lines do not represent observed data. Lines are meant as a guide to the eye.

- If few observed values a far apart or unevenly spaced, it is not suitable to connect points with lines.

]

.pull-right[
```{r, comment=NA, fig.cap='Figure 1', echo=FALSE}
library(mozzie)
library(ggplot2)
data(mozzie)
colombo.dengue <- mozzie[, 1:4]
ggplot(colombo.dengue, aes(x=ID, y=Colombo))+
  geom_point()+
  geom_line()+
  xlab("Time")+
  ylab("Number of dengue cases")
```

]

---

## Individual Time Series: Lines only

.pull-left[
```r
ggplot(colombo.dengue, aes(x=ID, y=Colombo))+
*  geom_line()+
  xlab("Time")+
  ylab("Number of dengue cases")
```

]

.pull-right[
```{r, comment=NA, fig.cap='Figure 1', echo=FALSE}
ggplot(colombo.dengue, aes(x=ID, y=Colombo))+
  geom_line()+
  xlab("Time")+
  ylab("Number of dengue cases")
```

]

---

## Individual Time Series: Lines only

.pull-left[

- Without points more emphasis is given on the overall trend and less on individual values.

- In general, when there are too many points it is better to plot without points.


]

.pull-right[
```{r, comment=NA, fig.cap='Figure 1', echo=FALSE}
ggplot(colombo.dengue, aes(x=ID, y=Colombo))+
  geom_line()+
  xlab("Time")+
  ylab("Number of dengue cases")
```

]

---

## Individual time series: Fill the area under the curve

.pull-left[
```r
ggplot(colombo.dengue, aes(x=ID, y=Colombo))+
*  geom_area()+
  xlab("Time")+
  ylab("Number of dengue cases")
```

]

.pull-right[
```{r, comment=NA, fig.cap='Figure 1', echo=FALSE}
ggplot(colombo.dengue, aes(x=ID, y=Colombo))+
  geom_area()+
  xlab("Time")+
  ylab("Number of dengue cases")
```

]

---

.pull-left[

- Visually separates the area above and below the curve.

- More emphasis is given to the overarching trend in the series. 

- This visualization is only valid if the y axis starts at zero.

]

.pull-right[
```{r, comment=NA, fig.cap='Figure 1', echo=FALSE}
ggplot(colombo.dengue, aes(x=ID, y=Colombo))+
  geom_area()+
  xlab("Time")+
  ylab("Number of dengue cases")
```

]

---

# Visualising multiple time series

```{r, echo=FALSE, comment=NA,message=FALSE, warning=FALSE}
library(tidyverse)
library(magrittr)
mozzie <- mozzie[,1:5]
district_count <- mozzie %>%
 pivot_longer(-c("ID", "Year", "Week"), names_to = "District", values_to = "count")
```

.pull-left[

```{r, comment=NA, echo=FALSE, fig.height=3}
ggplot(district_count, aes(x = ID, y = count, colour = District))+geom_point()

```

Difficult to read.

]

.pull-right[

```{r, comment=NA, echo=FALSE, fig.height=3}
ggplot(district_count, aes(x = ID, y = count, colour = District))+geom_line()

```

By connecting points with lines we help the reader to follow the paths of each individual time series.

]

---

# Smoothing

```{r}
library(gapminder)
head(gapminder)
```

---

# Smoothing

.pull-left[
```{r, comment=NA, message=FALSE, eval=FALSE}
library(tidyverse)
gapminder_af <- gapminder %>%
  filter(continent == "Africa")
ggplot(gapminder_af, aes(x=year, y=lifeExp))+
  geom_point()+
  geom_line()+
  facet_wrap(~country)
```

]



.pull-right[
```{r, comment=NA, message=FALSE, echo=FALSE}
library(tidyverse)
gapminder_af <- gapminder %>%
  filter(continent == "Africa")
ggplot(gapminder_af, aes(x=year, y=lifeExp))+
  geom_point()+
  geom_line()+
  facet_wrap(~country)
```

]

---
# Smoothing

.pull-left[
```{r, comment=NA, message=FALSE, eval=FALSE}
gapminder_af <- gapminder %>%
  filter(continent == "Africa")
ggplot(gapminder_af, aes(x=year, y=lifeExp))+
  geom_line()+
  facet_wrap(~country)+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

]



.pull-right[
```{r, comment=NA, message=FALSE, echo=FALSE}
gapminder_af <- gapminder %>%
  filter(continent == "Africa")
ggplot(gapminder_af, aes(x=year, y=lifeExp))+geom_line()+facet_wrap(~country)+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

]

---

# Smoothing


.pull-left[

```{r, comment=NA, message=FALSE, eval=FALSE}
library(tidyverse)
gapminder_af <- gapminder %>%
  filter(continent == "Africa")
ggplot(gapminder_af, aes(x=year, y=lifeExp))+geom_smooth()+facet_wrap(~country)+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```


]

.pull-right[

```{r, comment=NA, message=FALSE, echo=FALSE}
library(tidyverse)
gapminder_af <- gapminder %>%
  filter(continent == "Africa")
ggplot(gapminder_af, aes(x=year, y=lifeExp))+geom_smooth()+facet_wrap(~country)+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```


]

---

# Time series plots

.pull-left[

```{r, comment=NA, message=FALSE, warning=FALSE, fig.height=5, eval=FALSE}
library(forecast)
library(fpp2)
mozcol <- ts(mozzie$Colombo, frequency = 52, start = c(2008, 52))
autoplot(mozcol) +
  ggtitle("Dengue Count - Colombo") +
  ylab("Count") +
  xlab("Year")
```

]

.pull-right[

```{r, comment=NA, message=FALSE, warning=FALSE, fig.height=5, echo=FALSE}
library(forecast)
library(fpp2)
mozcol <- ts(mozzie$Colombo, frequency = 52, start = c(2008, 52))
autoplot(mozcol) +
  ggtitle("Dengue Count - Colombo") +
  ylab("Count") +
  xlab("Year")
```

]

---
# Seasonal plots

.pull-left[
```{r, comment=NA, message=FALSE, warning=FALSE, fig.height=5, eval=FALSE}
ggseasonplot(mozcol, year.labels=TRUE, year.labels.left=TRUE) +
  ylab("Count") +
  ggtitle("Seasonal plot: Dengue Count - Colombo")
```
]


.pull-right[
```{r, comment=NA, message=FALSE, warning=FALSE, fig.height=5, echo=FALSE}
ggseasonplot(mozcol, year.labels=TRUE, year.labels.left=TRUE) +
  ylab("Count") +
  ggtitle("Seasonal plot: Dengue Count - Colombo")
```
]
---
# Polar seasonal plot

.pull-left[

```{r, comment=NA, message=FALSE, warning=FALSE, fig.height=5, eval=FALSE}
ggseasonplot(mozcol, polar=TRUE) +
  ylab("Count") +
  ggtitle("Polar seasonal plot: Dengue Count - Colombo")
```

]


.pull-right[

```{r, comment=NA, message=FALSE, warning=FALSE, fig.height=5, echo=FALSE}
ggseasonplot(mozcol, polar=TRUE) +
  ylab("Count") +
  ggtitle("Polar seasonal plot: Dengue Count - Colombo")
```

]

---
# Time series plot

.pull-left[
```{r, comment=NA, message=FALSE, warning=FALSE, fig.height=5, echo=FALSE}
autoplot(a10) +
  ggtitle("Antidiabetic drug sales") +
  ylab("$ million") +
  xlab("Year")
```

]



.pull-right[
```{r, comment=NA, message=FALSE, warning=FALSE, fig.height=5, echo=FALSE}
autoplot(a10) +
  ggtitle("Antidiabetic drug sales") +
  ylab("$ million") +
  xlab("Year")
```

]
---
# Seasonal plot

.pull-left[

```{r, comment=NA, message=FALSE, warning=FALSE, fig.height=5, eval=FALSE}
ggseasonplot(a10, year.labels=TRUE, year.labels.left=TRUE) +
  ylab("$ million") +
  ggtitle("Seasonal plot: antidiabetic drug sales")
```

]


.pull-right[

```{r, comment=NA, message=FALSE, warning=FALSE, fig.height=5, echo=FALSE}
ggseasonplot(a10, year.labels=TRUE, year.labels.left=TRUE) +
  ylab("$ million") +
  ggtitle("Seasonal plot: antidiabetic drug sales")
```

]

---

# Polar seasonal plot

.pull-left[

```{r, comment=NA, message=FALSE, warning=FALSE, fig.height=5, eval=FALSE}
ggseasonplot(a10, polar=TRUE) +
  ylab("$ million") +
  ggtitle("Polar seasonal plot: antidiabetic drug sales")
```

]



.pull-right[

```{r, comment=NA, message=FALSE, warning=FALSE, fig.height=5, echo=FALSE}
ggseasonplot(a10, polar=TRUE) +
  ylab("$ million") +
  ggtitle("Polar seasonal plot: antidiabetic drug sales")
```

]
---

# Seasonal subseries plots

.pull-left[

```{r, comment=NA, message=FALSE, warning=FALSE, fig.height=5, eval=FALSE}
ggsubseriesplot(a10) +
  ylab("$ million") +
  ggtitle("Seasonal subseries plot: antidiabetic drug sales")
```

]


.pull-right[

```{r, comment=NA, message=FALSE, warning=FALSE, fig.height=5, echo=FALSE}
ggsubseriesplot(a10) +
  ylab("$ million") +
  ggtitle("Seasonal subseries plot: antidiabetic drug sales")
```

]

---

```{r, comment=NA}
a10
```
---

# Seasonal subseries plots

.pull-left[
```{r, comment=NA, message=FALSE, warning=FALSE, fig.height=5, echo=FALSE}
ggsubseriesplot(mozcol) +
  ylab("Count") +
  ggtitle("Seasonal subseries plot: Dengue count")
```

]

.pull-right[
```{r, comment=NA, message=FALSE, warning=FALSE, fig.height=5, echo=FALSE}
ggsubseriesplot(mozcol) +
  ylab("Count") +
  ggtitle("Seasonal subseries plot: Dengue count")
```

]

---
# Lag plots: a10

.pull-left[
Monthly anti-diabetic drug sales in Australia from 1991 to 2008.

```{r, comment=NA, message=FALSE, eval=FALSE}
gglagplot(a10)
```

]

.pull-right[
```{r, comment=NA, message=FALSE, echo=FALSE}
gglagplot(a10)
```
]
---

# Lag plots: mozzie 

.pull-left[
Dengue counts - Colombo

```{r, comment=NA, message=FALSE, eval=FALSE}
gglagplot(mozcol)
```
]

.pull-right[
```{r, comment=NA, message=FALSE, echo=FALSE}
gglagplot(mozcol)
```
]
---

# Lag plots: ausbeer

.pull-left[

Monthly Australian beer production: Jan 1991 – Aug 1995.

```{r, comment=NA, message=FALSE, eval=FALSE}
beer2 <- window(ausbeer, start=1992)
gglagplot(beer2)
```


]


.pull-right[


```{r, comment=NA, message=FALSE, echo=FALSE}
beer2 <- window(ausbeer, start=1992)
gglagplot(beer2)
```

]

---

### Time series features

Transform a given time series $y=\{y_1, y_2, \cdots, y_n\}$ to a feature vector $F = (f_1(y), f_2(y), \cdots, f_p(y))'$. 

**Examples of time series features**

- strength of trend

- strength of seasonality

- lag-1 autocorrelation

- spectral entropy

- proportion of zeros



---


.pull-left[


**Time-domain representation**

```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE}
library(Mcomp)
library(ggplot2)
library(grid)
library(gridExtra)
library(ggrepel)
library(png)
library(tsfeatures)
library(tidyverse)
library(ggpubr)
# Extract required series
series_id <- c("N0001", "N0633", "N0625", "N0645","N1912", "N2012")
color = c("#1b9e77", "#d95f02", "#7570b3", "#e7298a", "#66a61e", "#e6ab02")
six_series <- lapply(M3[series_id], function(u){u$x})
p <- lapply(six_series,
function(u) {autoplot(u) + xlab("") + ylab("")}
)
for (i in seq_along(six_series))
p[[i]] <- p[[i]] +geom_line(color=color[i]) +ggtitle(series_id[i])+theme(title =element_text(size=10, face='bold'))
grid.arrange(grobs = p, ncol = 2)
```

]


.pull-right[

**Feature-domain representation**

```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE}
df <- tsfeatures(six_series, c("stl_features")) %>%
  select(trend, seasonal_strength) %>%
  rename(seasonality = seasonal_strength) %>%
  replace_na(list(seasonality = 0))
df$id <- names(six_series)
ggplot(df, aes(x = trend, y = seasonality)) +
  geom_point(size = 5, color = c("#1b9e77", "#d95f02", "#7570b3", "#e7298a", "#66a61e", "#e6ab02")) +
  xlim(0, 1) + ylim(0, 1) +
  xlab("Strength of trend") + 
  ylab("Strength of seasonality") + 
  coord_fixed() +
  geom_text_repel(
    aes(label = id),
    colour = "black",
    size = 5,
    box.padding = unit(0.5, "lines")
  ) +
  theme(legend.position = "none")
```

]

---

## References



Talagala, T. S., Hyndman, R. J., & Athanasopoulos, G. (2018). Meta-learning how to forecast time series. Monash Econometrics and Business Statistics Working Papers, 6, 18.

