---
title: "Categorical Data "
author: "Thiyanga S. Talagala"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(skimr)
library(magrittr)
```


```{r, comment=NA, echo=FALSE}
skim(diamonds)
```

## Univariate

```{r, comment=NA}
p1 <- ggplot(data=diamonds, aes(x=cut))
```

1. R code: _______________________________

```{r, comment=NA, echo=FALSE}
p1 + geom_bar()
```

<!--The majorith of cuts are ideal, followed by Premium, with very few fair cuts.-->

2. R code: _______________________________

```{r, comment=NA, echo=FALSE}
p1 + geom_bar(fill="forestgreen")

```

3. R code: _______________________________

```{r, comment=NA, echo=FALSE}
p1 + geom_bar(fill="forestgreen")+
  labs(title="Composition of diamonds by cuts")
```

_____________

```{r, comment=NA}
#old version
#..count..: special variable to represent frequency
#p1 <- ggplot(data=diamonds, aes(x=cut, y=..count..))
#p2 <- ggplot(data=diamonds, aes(x=cut, #y=..count../sum(..count..)))

# New version
p1 <- ggplot(data=diamonds, aes(x=cut, y=after_stat(count/sum(count)))) + geom_bar()
p2 <- ggplot(data=diamonds, aes(x=cut, y=after_stat(count/sum(count)))) + geom_bar()
```


4. R code: _______________________________

```{r, comment=NA, echo=FALSE}
p2 + geom_bar(fill="forestgreen")+
  labs(title="Composition of diamonds by cuts")
```

5. R code: _______________________________

```{r, comment=NA, echo=FALSE}
p2 + geom_bar(fill="forestgreen")+
  labs(title="Composition of diamonds by cuts")+
  scale_y_continuous(labels = scales::percent)+
  labs(y="Percent")
```

 
6. R code: _______________________________

```{r, comment=NA, echo=FALSE}
cut.percent <- data.frame(cut=c("Fair", "Good", "Very Good", "Premium", "Ideal"),
      percent=c(3, 9, 22.4, 25.6, 40))
cut.percent
```


```{r, comment=NA}
p3 <- ggplot(data=cut.percent, aes(x=cut, y=percent))
```

```{r, comment=NA, echo=FALSE}
p3+geom_bar(stat="identity")
```

7. R code: _______________________________

```{r, comment=NA, echo=FALSE}
cut.percent$cut <- factor(cut.percent$cut, levels = c("Fair", "Good", "Very Good", "Premium", "Ideal"))

```

```{r, comment=NA}
# Need to rerun this once you change the factor levels (can't use p3)
ggplot(data=cut.percent, aes(x=cut, y=percent))+geom_bar(stat="identity")
```

8. Labeling bars: what is the suitable geom to add here?

```{r, comment=NA, echo=FALSE}
# Need to rerun this once you change the factor levels (can't use p3)
ggplot(data=cut.percent, aes(x=cut, y=percent))+geom_bar(stat="identity")+
  geom_text(aes(label = percent),
vjust=-0.5)
```

9. R code:___________

Help: use `coord_flip`

```{r, comment=NA, echo=FALSE}
p2 + geom_bar(fill="forestgreen")+
  labs(title="Composition of diamonds by cuts")+
  scale_y_continuous(labels = scales::percent)+
  labs(y="Percent")+
  coord_flip()
```

10. R Code:____________

```{r, comment=NA, echo=FALSE}
cut.prop <- cut.percent %>%
  mutate(prop=percent/100)
cut.prop
```


```{r, comment=NA}

ggplot(data=cut.prop, aes(x="", y=prop, fill=cut))+geom_bar(stat="identity", width=1)
```

```{r, comment=NA}

ggplot(data=cut.prop, aes(x="", y=prop, fill=cut))+geom_bar(stat="identity", width=1, position = "dodge")
```

11. R Code:____________

Pie charts are controversial in statistics.

<!--humans are better at judging the length of bars than the volume of pie slices).-->

```{r, comment=NA, echo=FALSE}

ggplot(data=cut.prop, aes(x="", y=prop, fill=cut))+geom_bar(stat="identity", width=1)+coord_polar("y",
start = 0,
direction = -1) 
```

Some extra work is needed to make the pie chart appealing to human eye.



<!--An alternative to a pie chart is a tree map. Unlike pie charts, it can handle categorical variables that have
many levels.-->
_______

## Bi-variate

### Stacked bar chart

**Encoding by colour**

Position: `stack`

```{r, comment=NA, message=FALSE}
b1 <- ggplot(data=diamonds, aes(x=cut, fill=color))
```

12: R code:___________

```{r, echo=FALSE}
#  stack is default
b1 + geom_bar(position = "stack")
```

### Grouped bar chart

**Encoding by colour**

Position: `dodge`

13: R code:___________

```{r, echo=FALSE}
#  stack is default
b1 + geom_bar(position = "dodge")
```

### Segmented bar chart

Position: `fill`

14: R code:___________

```{r, echo=FALSE}
#  stack is default
b1 + geom_bar(position = "fill")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
df <- data.frame(x1 = as.factor(c(1,1,1,1,1,1,1,1,0,0,0,0,0,0,0,0,0,0)),
                 x2= as.factor(c(1,1,0,0,0,1,1,1,0,0,0,0,0,0,1,1,0,0)))
ggplot(data=df, aes(x=x1, fill=x2))+
  geom_bar(position = "fill")
```

15: Rcode:_______________

**Encoding by position**

```{r, comment=NA, message==FALSE}
ggplot(data=diamonds, aes(x=color))+geom_bar()+facet_wrap(~cut)

```
----

## Categorical vs Quantitative

### Cleveland dot chart


<!--This is useful when you have large number of categories-->

```{r, echo=FALSE, comment=NA, message=FALSE, warning=FALSE}
library(gapminder)
gap2007 <- gapminder %>%
filter(continent == "Asia" &
year == 2007)

ggplot(gap2007,
aes(x= lifeExp, y = country)) +
geom_point()

ggplot(gap2007,
aes(x=lifeExp,
y=reorder(country, lifeExp))) +
geom_point()+ggtitle("life expectancy at birth, in 2007")

ggplot(gap2007,
aes(x=lifeExp,
y=reorder(country, lifeExp))) +
geom_bar(stat="identity")+ggtitle("life expectancy at birth, in 2007")

```

### Heat map

```{r comment=NA, message=FALSE, warning=FALSE, echo=FALSE}
gapminderAsia <- gapminder %>%
filter(continent == "Asia")

ggplot(gapminderAsia, aes(x=year, fill=lifeExp, y=country))+
  geom_raster()+
  scale_fill_viridis_c()

ggplot(gapminderAsia, aes(x=year, fill=lifeExp, y=reorder(country, lifeExp)))+
  geom_raster()+
  scale_fill_viridis_c()

ggplot(gapminderAsia, aes(x=year, y=lifeExp, col=country))+
  geom_line()
```
---

### Summary statistics

16. R code:___________________

```{r, comment=NA, message=FALSE, echo=FALSE}
cut.summary <- diamonds %>%
  group_by(cut) %>%
  summarize(mean_carat=mean(carat))
cut.summary
```

```{r, comment=NA, message=FALSE, echo=FALSE}
ggplot(cut.summary,
aes(x = cut,
y = mean_carat)) +
geom_bar(stat = "identity")

```

### Plotting summary statistics: `stat_summary`

17. R code:___________________

`mean_se`: mean and standard error

```{r, comment=NA, message=FALSE}
g1 <- ggplot(diamonds, aes(x = cut, y = carat)) 

```

```{r, comment=NA, message=FALSE, echo=FALSE}
ggplot(diamonds,
aes(x = cut, y = carat)) +
  geom_boxplot()

```

```{r, comment=NA, message=FALSE, echo=FALSE}
g1+stat_summary(fun.data = "mean_se")

```

18. R code:___________________

`mean_cl_normal`: 95 per cent confidence interval assuming normality. (Use `library(Hmisc)`)

```{r, comment=NA, message=FALSE, echo=FALSE}
library(Hmisc)
g1+stat_summary(fun.data = "mean_cl_normal")

```


19. R code:___________________


```{r, comment=NA, message=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
g1+stat_summary(fun.y = mean)

```


20. R code:___________________


```{r, comment=NA, message=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
g1+stat_summary(fun.y = summary)

```

21. R code:___________________


```{r, comment=NA, message=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
g1+stat_summary(fun.y = summary, alpha=0.5)+
  stat_summary(fun.y = "mean", geom="point", color="red")

```

22. R code:___________________


```{r, comment=NA, message=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
g1+stat_summary(fun.y = summary, alpha=0.5)+
  stat_summary(fun.y = "mean", geom="point", color="red")+
  stat_summary(fun.y = "median", geom="point", color="forestgreen")

```

23. R code:___________________

`mean_cl_boot`: Bootstrap confidence interval (95%)

```{r, comment=NA, message=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
g1 + stat_summary(fun.data = "mean_cl_boot") 

```

<!--Confidence limits may give us a better idea than standard error limits of whether two means
would be deemed statistically different when modeling, so we frequently use-->

24. R code:___________________

`mean_hilow`: Median, Q1, Q3


```{r, comment=NA, message=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
g1 + stat_summary(fun.data = "mean_cl_boot") 

```

## Design of Experiments

Description

The response is the length of odontoblasts (cells responsible for tooth growth) in 60 guinea pigs. Each animal received one of three dose levels of vitamin C (0.5, 1, and 2 mg/day) by one of two delivery methods, orange juice or ascorbic acid (a form of vitamin C and coded as VC).

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=NA}
data(ToothGrowth)
skim(ToothGrowth)
head(ToothGrowth)
```

25. R code:___________________

```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE}
ggplot(ToothGrowth, aes(x = dose, y = len, colour = supp)) +
stat_summary(fun.data = "mean_se")
```

26. R code:___________________

```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE}
ggplot(ToothGrowth, aes(x = dose, y = len, colour = supp)) +
stat_summary(fun.data = "mean_se")+
  stat_summary(fun.y = mean, geom = "line")
```

27. R code:____________________

Avoid overlapping in the last category `position_dodge(0.1)`

```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE}
ggplot(ToothGrowth, aes(x = dose, y = len, colour = supp)) +
stat_summary(fun.data = "mean_se", position= position_dodge(0.1))+
stat_summary(fun.y = mean, geom = "line", position= position_dodge(0.1))
```

28. R code: ___________

Not suitable for this example: Why?

```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE}
ggplot(ToothGrowth, aes(x = dose, y = len, colour = supp, fill = supp)) +
stat_summary(fun.y = mean, geom = "line") +
stat_summary(fun.data = mean_se, geom = "ribbon", alpha = .5)
```

## Categorical with two Quantitative variables

29. R code: ___________


```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE}
ggplot(diamonds, aes(y = price,
x = carat,
color=color)) +
geom_point()

```

<!--Price colour relasionship changes according to the colour-->

30. R code: ___________

```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE}
ggplot(diamonds, aes(y = price,
x = carat,
color=color)) +
geom_point(alpha=0.5)

```

31. R code: ___________

```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE}
ggplot(diamonds, aes(y = price,
x = carat,
color=color)) +
geom_point(alpha=0.5)+
  geom_smooth(se=FALSE)

```

32. R code: ___________

```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE}
ggplot(diamonds, aes(y = price,
x = carat,
color=color)) +
  geom_smooth(se=FALSE)

```

33. R code: ___________

```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE}
ggplot(diamonds, aes(y = price,
x = carat,
color=color)) +
  geom_smooth(se=TRUE)

```

34. R code: ___________

```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE}
ggplot(diamonds, aes(y = price,
x = carat,
color=color)) +
  geom_smooth(se=FALSE)+facet_grid(~color)

```
