---
output: 
  html_document:
    toc: true
    toc_float: true 
    theme: darkly
---

# Practical Class Tutorial: Week 4

Learning objective:

- Produce barcharts, scatter plots, and boxplots plots using the ggplot2 package.

- Apply faceting in ggplot.


- Customized plots using `scale_*`

# Resources

ggplot cheat sheet: [Click here](/slides/data-visualization-2.1.pdf)

# Data set

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This question is based on `diamonds` dataset in the `ggplot2` package. Load the `ggplot2` package and `diamonds` dataset using the following command.

```{r, comment=NA}
library(ggplot2)
data(diamonds)
```


You can observe the data structure using the following command

```{r, comment=NA}
str(diamonds)
```

To get a summary use the command

```{r, comment=NA}
summary(diamonds)
```

# Exercise

## Write R codes to reproduce the following plots using  the `ggplot` function in the `ggplot2` package. 

### Plot 1

```{r, comment=NA, echo=FALSE}
ggplot(diamonds, aes(x=cut)) + geom_bar()

```

### Plot 2

**Help: `?geom_bar`**

**Click here to view the colour names:** https://www.r-graph-gallery.com/42-colors-names.html

```{r, comment=NA, echo=FALSE}
ggplot(diamonds, aes(x=cut)) + geom_bar(colour="cornflowerblue")

```

### Plot 3

```{r, comment=NA, echo=FALSE}
ggplot(diamonds, aes(x=cut)) + geom_bar(colour="cornflowerblue", fill="deeppink")

```

### Plot 4

```{r, comment=NA, echo=FALSE}
ggplot(diamonds, aes(x=cut)) + geom_bar(colour="cornflowerblue", fill="cornflowerblue")

```

### Plot 5

```{r, comment=NA, echo=FALSE}
ggplot(diamonds, aes(x=cut)) + geom_bar(aes(fill=cut))

```

### Plot 6

```{r, comment=NA, echo=FALSE}
ggplot(diamonds, aes(x=cut, fill=cut)) + geom_bar(aes( y=after_stat(100*count/sum(count))))

```

### Plot 7

```{r, comment=NA, echo=FALSE}
ggplot(diamonds, aes(carat, fill = cut)) +
  geom_density(position = "stack")
```

### Plot 8

```{r, comment=NA, echo=FALSE}
ggplot(diamonds, aes(depth, fill = cut, colour = cut)) +
  geom_density(alpha = 0.1)
```

### Plot 9

```{r, comment=NA, echo=FALSE}
ggplot(diamonds, aes(depth, colour = cut)) +
  geom_density()
```

### Plot 10

```{r, comment=NA, echo=FALSE}
ggplot(diamonds, aes(x = color, y = carat)) +
  geom_boxplot(aes(fill = color))
```


### Plot 11

```{r, comment=NA, echo=FALSE}
ggplot(diamonds, aes(x = color, y = carat)) +
  geom_boxplot(aes(fill = color)) +
  facet_wrap(~cut, scales="free")

```

### Plot 12

```{r, comment=NA, echo=FALSE}
ggplot(diamonds, aes(y=price, x=carat)) +
  geom_point()
```


### Plot 13

```{r, comment=NA, echo=FALSE}
ggplot(diamonds, aes(y=price, x=carat, color=cut)) +
  geom_point()
```

### Plot 14

**Help (R colorbrewer palette):** https://www.datanovia.com/en/blog/the-a-z-of-rcolorbrewer-palette/

```{r, comment=NA, echo=FALSE}
ggplot(diamonds, aes(y=price, x=carat, color=cut)) +
  geom_point() + scale_color_brewer(type="seq", palette="BuPu")
```


### Plot 15


```{r, comment=NA, echo=FALSE}
ggplot(diamonds, aes(y=price, x=carat, color=cut)) + geom_point() + 
  scale_color_brewer(type="div", palette="Dark2")
```

### Plot 16


```{r, comment=NA, echo=FALSE}
ggplot(diamonds, aes(y=price, x=carat, color=cut)) + geom_smooth() 
```

### Plot 17


```{r, comment=NA, echo=FALSE}
ggplot(diamonds, aes(y=price, x=carat, color=cut)) + geom_point() + geom_smooth() 
```

### Plot 18

```{r, comment=NA, echo=FALSE}
ggplot(diamonds, aes(y=price,  x=cut)) + geom_jitter() 
```

### Plot 19

```{r, comment=NA, echo=FALSE}
ggplot(diamonds, aes(y=price,  x=cut)) + geom_jitter(alpha=0.5) 
```

### Plot 20

```{r, comment=NA, echo=FALSE}
ggplot(diamonds, aes(y=price,  x=cut)) + geom_violin() 
```

