---
title: "ASP 460 2.0 Data Visualisation"
subtitle: "Higher-dimensional Displays and Special Structures"
author: "Thiyanga Talagala"
date: ""
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: 
      - default
      - default-fonts
      - duke-blue
      - hygge-duke
      - libs/cc-fonts.css
      - libs/figure-captions.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```


## Higher-dimensional Displays and Special Structures

- Scatterplot Matrices (Sploms) 

- Parallel Coordinates 

- Mosaic Plots 

- Small Multiples and Trellis Displays 

- Time Series

---
## Scatterplot Matrices

```{r, comment=NA, message=FALSE, echo=FALSE}
#install.packages("ggplot2")
#install.packages("GGally")
library(GGally)
library(ggplot2)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(ggplot2)
library(cdata)
meas_vars <- colnames(iris)[1:4]
controlTable <- data.frame(expand.grid(meas_vars, meas_vars,                               stringsAsFactors = FALSE))
# rename the columns
colnames(controlTable) <- c("x", "y")

# add the key column
controlTable <- cbind(
  data.frame(pair_key = paste(controlTable[[1]], controlTable[[2]]),
             stringsAsFactors = FALSE),
  controlTable)
iris_aug = rowrecs_to_blocks(
  iris,
  controlTable,
  columnsToCopy = "Species")
splt <- strsplit(iris_aug$pair_key, split = " ", fixed = TRUE)
iris_aug$xv <- vapply(splt, function(si) si[[1]], character(1))
iris_aug$yv <- vapply(splt, function(si) si[[2]], character(1))

iris_aug$xv <- factor(as.character(iris_aug$xv),
                           meas_vars)
iris_aug$yv <- factor(as.character(iris_aug$yv),
                           meas_vars)


ggplot(iris_aug, aes(x=x, y=y)) +
  geom_point(aes(color=Species, shape=Species)) + 
  facet_grid(yv~xv, labeller = label_both, scale = "free") +
  ggtitle("Anderson's Iris Data -- 3 species") +
  scale_color_brewer(palette = "Dark2") +
  ylab(NULL) + 
  xlab(NULL)
                                       
```

---


```{r, comment=NA, echo=FALSE, message=FALSE, warning=FALSE}
library(GGally)
ggpairs(iris)

```

---

```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE}
ggpairs(iris, aes(colour = Species, alpha = 0.4))
```

---


```{r, comment=NA}
set.seed(2020)
x <- as.factor(c(rep(1, 100), rep(0, 100)))
y <- as.factor(sample(x))
z <- as.factor(c(rep(1, 25), rep(2, 75), rep(3, 50), rep(4, 50)))
x <- factor(x, label=c("A", "B"))
y <- factor(y, label=c("Yes", "No"))
z <- factor(z, label=c("High", "Middle", "Low", "Never"))
df <- data.frame(v1=x, v2=y, v3=z)
summary(df)

```

---


```{r, comment=NA, message=FALSE}
table(df$v1,df$v2)
table(df$v1, df$v3)
table(df$v2, df$v3)

```

---

# Mosaic Plot (1 ~ X)

.pull-left[

```{r, comment=NA, message=FALSE, echo=FALSE}
table(df$v1,df$v2)
table(df$v1, df$v3)
table(df$v2, df$v3)

```

]

.pull-right[

```{r, comment=NA, message=FALSE}
library(ggmosaic)
ggplot(df)+
  geom_mosaic(aes(x=product(v1)))+
  ggtitle("V1")
```

]
---


# Mosaic Plot (1 ~ X)

.pull-left[

```{r, comment=NA, message=FALSE, echo=FALSE}
summary(df)
table(df$v1,df$v2)
table(df$v1, df$v3)
table(df$v2, df$v3)

```

]

.pull-right[

```{r, comment=NA, message=FALSE}
ggplot(df)+
  geom_mosaic(aes(x=product(v1), 
                  fill=v1))+
  ggtitle("V1")
```

]

---


# Mosaic Plot (1 ~ X)

.pull-left[

```{r, comment=NA, message=FALSE, echo=FALSE}
summary(df)
table(df$v1,df$v2)
table(df$v1, df$v3)
table(df$v2, df$v3)

```

]

.pull-right[

```{r, comment=NA, message=FALSE}
ggplot(df)+
  geom_mosaic(aes(x=product(v3), 
                  fill=v3))+
  ggtitle("V3")
```

]

---

# Mosaic Plot (1 ~ Y + X)


.pull-left[

```{r, comment=NA, message=FALSE, echo=FALSE}
summary(df)
table(df$v1,df$v2)
table(df$v1, df$v3)
table(df$v2, df$v3)

```

]

.pull-right[

```{r, comment=NA, message=FALSE}
ggplot(df)+
geom_mosaic(aes(x=product(v1, v2),
fill=v2))+ggtitle("V1 and V2*")
```

]

---

# Mosaic Plot (1 ~ Y + X)


.pull-left[

```{r, comment=NA, message=FALSE, echo=FALSE}
summary(df)
table(df$v1,df$v2)
table(df$v1, df$v3)
table(df$v2, df$v3)

```

]

.pull-right[

```{r, comment=NA, message=FALSE}
ggplot(df)+
geom_mosaic(aes(x=product(v1, v2),
fill=v1))+ggtitle("V1* and V2")
```

]

---


# Mosaic Plot (1 ~ Y + X)


.pull-left[

```{r, comment=NA, message=FALSE, echo=FALSE}
summary(df)
table(df$v1,df$v2)
table(df$v1, df$v3)
table(df$v2, df$v3)

```

]

.pull-right[

```{r, comment=NA, message=FALSE}
ggplot(df)+
geom_mosaic(aes(x=product(v1, v3),
fill=v1))+ggtitle("V1 and V3*")
```

]
---


# Mosaic Plot (1 ~ Y + X)


.pull-left[

```{r, comment=NA, message=FALSE, echo=FALSE}
summary(df)
table(df$v1,df$v2)
table(df$v1, df$v3)
table(df$v2, df$v3)

```

]

.pull-right[

```{r, comment=NA, message=FALSE}
ggplot(df)+
geom_mosaic(aes(x=product(v3, v1),
fill=v1))+ggtitle("V1 and V3*")
```

]

---


# Mosaic Plot (1 ~ Y + X)


.pull-left[

```{r, comment=NA, message=FALSE, echo=FALSE}
summary(df)
table(df$v1,df$v2)
table(df$v1, df$v3)
table(df$v2, df$v3)

```

]

.pull-right[

```{r, comment=NA, message=FALSE}
ggplot(df)+
geom_mosaic(aes(x=product(v3, v1),
fill=v3))+ggtitle("V1 and V3*")
```

]

---


# Mosaic Plot (1 ~ Y + X|Z)


.pull-left[

```{r, comment=NA, message=FALSE, echo=FALSE}
library(tidyverse)
dfv2yes <- filter(df, v2=="Yes")
dfv2no <- filter(df, v2=="No")
table(dfv2yes$v1, dfv2yes$v3)
table(dfv2no$v1, dfv2no$v3)
```

]

.pull-right[

```{r, comment=NA, message=FALSE}
ggplot(df)+
geom_mosaic(aes(x=product(v3, v1),
fill=v3))+ggtitle("V1 and V3*") + 
  facet_wrap(~v2)
```

]



---
# Conditioning plots (Coplots)

## Trellis Graphics

- Grid graphic system

- The idea is based on conditioning on the
values taken on by one or more of the variables in a data
set.

    - Categorical: same subplot for levels of the conditioning variable
    
    - Numeric: same subplot for different ranges (intervals) of the conditioning variable

---
# Trellis Graphics: Categorical

```{r, comment=NA, message=FALSE, echo=FALSE}
ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width))+
  geom_point() + facet_wrap(~Species)+coord_fixed(ratio=2)

```

---
# Trellis Graphics: Numeric

`quakes`: magitude of earthquakes under the Tonga
Trench, to the North of New Zealand.

```{r, comment=NA, message=FALSE, warning=FALSE}
str(quakes)
```

```{r, comment=NA, message=FALSE, warning=FALSE}
library(lattice)
```
---

# Trellis Graphics: Numeric


```{r, echo=FALSE}
library(lattice)
Depth = equal.count(quakes$depth,
number = 6,
overlap = .1)

xyplot(lat ~ long | Depth, data = quakes,
xlab = "Longtitude",
ylab = "Lattitude")

```

---

.pull-left[

## Unconditional plot

```{r, comment=NA, echo=FALSE, warning=FALSE, message=FALSE}
xyplot(lat ~ long, data = quakes,
xlab = "Longtitude",
ylab = "Lattitude")
```

]

.pull-right[

## Conditional plot


```{r, comment=NA, echo=FALSE, warning=FALSE, message=FALSE}
Depth = equal.count(quakes$depth,
number = 6,
overlap = .1)

xyplot(lat ~ long | Depth, data = quakes,
xlab = "Longtitude",
ylab = "Lattitude")
```

Condition on `depth`
]

---

# Coplots using ggplot

.pull-left[

```{r, comment=NA, echo=FALSE, warning=FALSE, message=FALSE}
Depth = equal.count(quakes$depth,
number = 6,
overlap = .1)

xyplot(lat ~ long | Depth, data = quakes,
xlab = "Longtitude",
ylab = "Lattitude")
```

]

.pull-right[
```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE}
ggplot(quakes, aes(x=long, y=lat)) + 
geom_point() + facet_wrap(~ cut_number(depth, 6))
```
]
---


# Coplots using ggplot

.pull-left[

```{r, comment=NA, echo=FALSE, warning=FALSE, message=FALSE}
Depth = equal.count(quakes$depth,
number = 6,
overlap = .1)

xyplot(lat ~ long | Depth, data = quakes,
xlab = "Longtitude",
ylab = "Lattitude")
```

]

.pull-right[
```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE}
ggplot(quakes, aes(x=long, y=lat)) + 
geom_point() + facet_wrap(~ cut_number(depth, 6)) + geom_smooth()
```
]
---
# Parallel Coordinates Plots

```{r, comment=NA}
df1 <- data.frame(x=c(1, 2), y=c(20, 10), z=c(10, 10))
df1

```


```{r, comment=NA, fig.height=3, echo=FALSE}
ggparcoord(df1)

```

---

```{r, comment=NA, echo=FALSE}
ggparcoord(data = iris)
```

---
```{r, comment=NA, echo=FALSE}
ggparcoord(data = iris[,1:4])
```


---
```{r, comment=NA, echo=FALSE}
ggparcoord(data = iris, columns = 1:4, groupColumn = 5)
```
---

```{r, comment=NA, echo=FALSE}

ggparcoord(data = iris, columns = 1:4, groupColumn = 5, order = "anyClass")
```

---
```{r, comment=NA, echo=FALSE}

ggparcoord(data = iris, columns = 1:4, groupColumn = 5, order = "anyClass", alphaLines = 0.5)
```

---
## ggparcoord: scales

The different types of scales are as follows:

`std`: univariately, subtract mean and divide by standard deviation

`robust`: univariately, subtract median and divide by median absolute deviation

`uniminmax`: univariately, scale so the minimum of the variable is zero, and the maximum is one

`globalminmax`: no scaling is done; the range of the graphs is defined by the global minimum and the global maximum

`center`: use uniminmax to standardize vertical height, then center each variable at a value specified by the scaleSummary param

`centerObs`: use uniminmax to standardize vertical height, then center each variable at the value of the observation specified by the centerObsID param
---

```{r, comment=NA, echo=FALSE, message=FALSE}
library(gridExtra)
g1<-ggparcoord(df1, columns=1:3, scale = "std", title = "Standard Scale")
g2<-ggparcoord(df1, columns=1:3, scale = "robust", title = "Robust Scale")
g3<-ggparcoord(df1, columns=1:3, scale = "uniminmax", title = "Uniminmax Scale")
g4<-ggparcoord(df1, columns=1:3, scale = "globalminmax", title = "Globalminmax Scale")
g5<-ggparcoord(df1, columns=1:3, scale = "center", scaleSummary = "mean", title = "Center Scale")
g6<-ggparcoord(df1, columns=1:3, scale = "centerObs", centerObsID = 1, title = "CenterObs Scale")

grid.arrange(g1, g2, g3, g4, g5, g6, nrow=2)
```