---
title: "Data Visualization"
subtitle: "Introduction to Time Series"
author: "Dr Thiyanga Talagala"
date: "2020-03-12"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: 
      - default
      - default-fonts
      - duke-blue
      - hygge-duke
      - libs/cc-fonts.css
      - libs/figure-captions.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```


## Time series

- A time series is a sequence of observations taken sequentially in time. 

## Time series data vs Cross sectional data

.pull-left[

- Time series data

```{r, comment=NA, echo=FALSE}
library(knitr)
library(magrittr)
library(kableExtra)
ts <- data.frame(Year=c(2012, 2013, 2014, 2015),
                 Values =c(120, 122, 140, 150))
kable(ts, format="html", align = "l", booktabs = T, escape = FALSE)
```

- a set of observations, along with some information about what times those observations were recorded.

- usually discrete and equally spaced time intervals.

]

.pull-right[

- Cross-sectional data

```{r, comment=NA, echo=FALSE}
library(knitr)
library(magrittr)
library(kableExtra)
ts <- data.frame(ID=c(1, 2, 3, 4),
                 Values =c(200, 350, 480, 250))
kable(ts, format="html", align = "l", booktabs = T, escape = FALSE)
```

- observations that come from different individuals or groups at a single point in time.

]
---

## Deterministic  vs Non-deterministic time series

.pull-left[
- **Deterministic time series:** future values can be exactly determined by using some mathematical function.



```{r, echo=FALSE, comment=NA, message=FALSE}
library(ggplot2)
library(forecast)
x <- 0:5000
y <- cos(2*3.14*x)
y <- ts(y, frequency=1)
autoplot(y)

```


$$y_t = cos(2\pi t)$$

]

.pull-right[
- **Non-deterministic time series:** future values can be determined only in terms of a probability distribution.



```{r, echo=FALSE, comment=NA, message=FALSE}
library(mozzie)
y <- ts(mozzie$Colombo)
#y <- ts(mozzie$Colombo, frequency=52, start=c(2008, 51))
autoplot(y)+ylab("Number of dengue cases in Colombo")

```

]
---

## Frequency of a time series: Seasonal periods

- Frequency: number of observation per natural time interval of measurement (usually year,  but sometimes a week, a day or an hour)

```{r, comment=NA, echo=FALSE}
library(knitr)
library(magrittr)
library(kableExtra)
ts <- data.frame(Data=c("Annual", "Quarterly", "Monthly", "Weekly"),
                  Frequency =c("1", "4", "12", "52 or 52.18"))
kable(ts, format="html", align = "l", booktabs = T, escape = FALSE)
```

- Multiple frequency setting

```{r, comment=NA, echo=FALSE}
library(knitr)
library(magrittr)
library(kableExtra)
tsf <- data.frame(Data=c("Daily", "Hourly", "Half-Hourly", "Minutes", "Seconds"), Minute=c("", "", "", "", "60"),
                 Hour=c("", "", "", "60", "3600"),
                 Day=c("", "24", "48", "1440", "86400"),
                 Week=c("7", "168", "336", "10080", "604800"),
                 Year=c("365.25", "8766", "17532", "525960", "31557600")
                 )
kable(tsf, format="html", align = "l", booktabs = T, escape = FALSE)
```  



---
.pull-left[

## Monthly time series

```{r, echo=FALSE}

y <- ts(USAccDeaths, start=1949, frequency = 12)
autoplot(y)
```

- Length of the series: 72

- Monthly seasonality

]


.pull-right[

## Half-hourly Time Series

```{r, echo=FALSE}
x <- msts(taylor, seasonal.periods=c(48,336), start=2000+22/52)
autoplot(x)
```

- Length of the series: 4032

- Daily seasonality and weekly seasonality

]

--
Note: Monthly seasonality with high-frequency data (daily, hourly, etc.) is tricky due to variable month lengths. You can't specify that using seasonal periods. It could be possibly handled using a dummy variable. 

---

class: duke-orange

# Your turn

- What are the frequencies for a monthly time series with semi-annual and annual pattern?


---

# `ts` object in R

- Annual time series

```{r, comment=NA, echo=FALSE}
library(knitr)
library(magrittr)
library(kableExtra)
tsdataset <- data.frame(Year=c(2012, 2013, 2014, 2015),
                 Values =c(120, 122, 140, 150))
kable(tsdataset, format="html", align = "l", booktabs = T, escape = FALSE)
```



```{r, comment=NA}
y <- ts(c(120, 122, 140, 150), start=2012)
y
```

---

# `ts` object in R

- Quarterly time series

```{r, comment=NA, echo=FALSE}
library(knitr)
library(magrittr)
library(kableExtra)
tsdataset <- data.frame(Quarter=c("2012-Q1", "2012-Q2", "2012-Q3", "2012-Q4", "2013-Q1"),
                 Values =c(120, 122, 140, 150, 200))
kable(tsdataset, format="html", align = "l", booktabs = T, escape = FALSE)
```



```{r, comment=NA}
y <- ts(c(120, 122, 140, 150, 200), start=c(2012, 1), frequency = 4)
y
```



---
# `ts` object in R

- Monthly time series

```{r, comment=NA, echo=FALSE}
library(knitr)
library(magrittr)
library(kableExtra)
tsdataset <- data.frame(Month=c("2012-Jan", "2012-Feb", "2012-March", "2012-April"),
                 Values =c(120, 122, 140, 150))
kable(tsdataset, format="html", align = "l", booktabs = T, escape = FALSE)
```



```{r, comment=NA}
y <- ts(c(120, 122, 140, 150), start=c(2012, 1), frequency = 12)
y
```

---
# `ts` object in R

- Weekly time series

```{r, comment=NA, echo=FALSE}
library(knitr)
library(magrittr)
library(kableExtra)
tsdataset <- data.frame(Week=c("2012-W1", "2012-W2", "2012-W3", "2012-W4"),
                 Values =c(120, 122, 140, 150))
kable(tsdataset, format="html", align = "l", booktabs = T, escape = FALSE)
```



```{r, comment=NA}
y <- ts(c(120, 122, 140, 150), start=c(2012, 1), frequency = 52)
y
```

---

## Time series plots

.pull-left[

```{r, comment=NA, echo=FALSE}
library(knitr)
library(magrittr)
library(kableExtra)
tsdataset <- data.frame(Year=c(2012, 2013, 2014, 2015, 2016, 2017),
                 Values =c(120, 122, 140, 150, 200, 250))
kable(tsdataset, format="html", align = "l", booktabs = T, escape = FALSE)
```

```{r, comment=NA}
y <- ts(c(120, 122, 140, 150, 200, 250), start=2012)
y
class(y)
```

]


.pull-right[

```{r, comment=NA}
autoplot(y)
```


]

---

## Add title and labels 

.pull-left[

```{r, comment=NA}
autoplot(y)
```

]

.pull-right[

```{r, comment=NA}
autoplot(y)+ylab("Number of sales")+
  xlab("Year")+
  ggtitle("Time series plot of sales from 2012 to 2017")
```

]

---

class: duke-orange

## Your turn

Create plots of the following time series: dengue counts in Gampaha (Use `mozzie` package), a10 series (`fpp2` package). 

Use help() to find out about the data in each series.

Modify the axes labels and title.

---

#### Time series patterns

**Trend**

- Long-term increase or decrease in the data.

**Seasonal**


- A seasonal pattern exists when a series is influenced by seasonal factors (e.g.,
the quarter of the year, the month, or day of the week). Seasonality is always
of a **fixed** and **known period**. Hence, seasonal time series are sometimes called
periodic time series.

- Period is unchanging and associated with some aspect of the
calendar.


**Cyclic**

- A cyclic pattern exists when data exhibit rises and falls that are not of fixed
period. The duration of these fluctuations is usually of at least 2 years.



In general, 

  - the average length of cycles is longer than the length of a seasonal pattern.

  - the magnitude of cycles tends to be more variable than the magnitude of seasonal patterns.
  
---

## Cyclic pattern

```{r, echo=FALSE, comment=NA, message=F}
library(fpp2)
autoplot(lynx) + xlab("Year") + ylab("Number of lynx trapped")

```
  
---

## Cyclic and seasonal pattern

```{r, echo=FALSE, comment=NA}
library(fpp2)
autoplot(hsales) + xlab("Year") + ylab("Monthly housing sales
(millions)")

```

---


## Multiple seasonal pattern

```{r, echo=FALSE, comment=NA}
library(fpp2)
autoplot(taylor) + xlab("Week") + ylab("Electricity demand (GW)")

```
  
---

## Seasonal pattern

```{r, echo=FALSE, comment=NA}
beer <- window(ausbeer,start=1992)
autoplot(beer)
```  
---

## Trend


```{r, echo=FALSE, comment=NA}
library(fpp2)
x <- 1:20
y <- 5*x+rnorm(20)
y <- ts(y, frequency = 1, start=1980)
autoplot(y)

```

---

## Trend and seasonal

```{r, comment=NA, echo=FALSE}
autoplot(a10) + ylab("$ million") + xlab("Year") +
ggtitle("Antidiabetic drug sales")
```

