---
title: "ASP 460 2.0 Data Visualization"
author: "Dr Thiyanga Talagala"
date: "Principal Component Analysis"
output:
  xaringan::moon_reader:
    chakra: libs/remark-latest.min.js
    lib_dir: libs
    css: 
      - default
      - default-fonts
      - duke-blue
      - hygge-duke
      - libs/cc-fonts.css
      - libs/figure-captions.css
      - xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
---
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent(
  base_color = "#081d58",
  text_bold_color = "#ce1256",
  title_slide_text_color = "#edf8b1",
  header_font_google = google_font("Josefin Sans"),
  base_font_size = "20px",
  text_font_size = "1.5rem"
 #text_font_google   = google_font("Montserrat", "300", "300i")
 # code_font_google   = google_font("Fira Mono")
)
```

# PCA

- Finding low-dimensional combinations (or projections) of high dimensional data that capture most of the variability in the original data.

- Objective: Take $P$ variables $X_1, X_2, X_3,...X_p$ and find combinations of these to produce variables (components) $Z_1, Z_2...Z_p$ that are uncorrelated.

> Number of PCs =  Number of variables in the original data.

- The components are ordered so that $Z_1$ captures the largest proportion of the data, $Z_2$ captures the next largest proportion of the variability

> $Var(Z_1) \geq Var(Z_2)...\geq Var(Z_p)$

---

# PCA (cont.)


- If original variables are uncorrelated then PCA does nothing at all. The PCs are the same as the original data.

- PCs are formed by calculating the eigen vectors and eigen values of the data covariance matrix.

---


