---
title: "ASP 460 2.0 Special topics in Statistics: Data Wrangling"
date: "13/05/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Packages 

## Install tidyverse into your computer.

```r
install.packages("tidyverse")
```

## We are going to work with datasets in `EDAWR`.

Installation of EDAWR is bit different. Use the following command.

**Step 1**

```r
install.packages("devtools")
```

**Step 2**

```r
devtools::install_github("rstudio/EDAWR")
```

## Load packages

```{r}
library(tidyverse)
library(EDAWR) # load data
```

# Pipe operator, dplyr, and tidyr


- `dplyr` is a package for data wrangling, with several key verbs (functions).

- `slice()` and `filter()`: subset rows based on numbers or conditions.

- `select()` and `pull()`: select columns or a single column as a vector.

- `arrange()`: order rows by one or multiple columns.

- `rename()` and `mutate()`: rename or create columns.

- `mutate_at()`: apply a function to given columns.

Recall: Pipe operator (`%>%`)
===

See the slides in STA 326 2.0

Link: : https://hellor.netlify.app/slides/l7_intro_tidyverse.html#43 

```{r, message=FALSE, warning =F, comment=NA}
iris %>% 
  filter(Sepal.Length >= 7) 

iris %>% 
  filter(Sepal.Length >= 7) %>% 
  head(2)
```


## `tidyr` functions (They are also called tidy R verbs)

Main verbs (functions) in `tidyr`:

- `pivot_longer()`: makes datasets longer by increasing the number of rows and decreasing the number of columns. 

- `pivot_wider()`: is the opposite of pivot_longer() : it makes a dataset wider by increasing the number of columns and decreasing the number of rows.

- `separate()`: splits a single column into multiple columns.

- `unite()`: combines multiple columns into a single column.


`pivot_longer()`
===


```{r warning=F, message=F}
# EDAWR::cases means cases dataset in EDAWR
EDAWR::cases %>% 
  head(3)

EDAWR::cases %>% 
  pivot_longer(names_to = "year", values_to = "n", cols = 2:4) %>% 
  head(5)
```


- Here the columns 2:4 are transposed into a `year` column.
- We put the corresponding count values into a column called `n`.

## Other approaches to do the same thing.

```{r}
# Method 2
EDAWR::cases %>% 
  pivot_longer(names_to = "year", values_to = "n", -country) %>% 
  head(5)
# Method 3
# EDAWR::cases %>% 
#   pivot_longer(names_to = "year", values_to = "n", c(`2011`, `2012`, `2013`)) 
# Method 4
# EDAWR::cases %>% 
#   pivot_longer(names_to = "year", values_to = "n",  `2011`:`2013`) 
```

`pivot_wider()`: Makes longer data formats wider.
===


```{r}
EDAWR::pollution %>% 
  head(5)

EDAWR::pollution %>% 
  pivot_wider(names_from = "size", 
              values_from = "amount")
```



When could I use these operations?
===

- Data visualization with `ggplot2`.

**Read**:  https://tidyr.tidyverse.org/articles/pivot.html#manual-specs

## `separate()` 

To separate a character column into multiple columns using a regular expression separator.

The following code seperates date into multiple columns. "-" is used to seperate between words.

```{r, cache=TRUE}
EDAWR::storms %>% 
  head(3)

storms2 <- EDAWR::storms %>% 
  separate(date, c("y", "m", "d"), sep="-") # sep = "-"
storms2
```

`unite()`
===

Paste together multiple columns into one. 

The following code combines `y`, `m` and `d` in `storm2` using "-".

```{r}
storms2 %>%
  unite(date, y, m, d, sep = "-")
```



Note that `unite()` and `separate()` are inverse operations.


# `dplyr` package

`group_by()`
===

To define a grouping of rows based on a column:

```{r}
iris %>% 
  group_by(Species) %>%
  head(4)

iris %>% 
  group_by(Species) %>%
  head(4) %>% class
```


- This doesn't actually change anything in the output.

- The only difference is that when it prints, we're told about the groups.

- But it will play a big role in how other `dplyr` functions work. 

`summarize()` (American) or `summarise()` (British)
===

`summarize()` or `summarise()` in `dplyr` gives you single numerical summaries.

```{r}
# Ungrouped
iris %>% 
  summarize(Sepal.Length = mean(Sepal.Length),
            Sepal.Width = mean(Sepal.Width))

# Grouped by number of Species
iris %>%
  group_by(Species) %>%
  summarize(Sepal.Length = mean(Sepal.Length),
            Sepal.Width = mean(Sepal.Width))
```

---

```{r}
iris %>%
  group_by(Species) %>%
  summarize(Sepal.Width_mean = mean(Sepal.Width),
            Sepal.Width_max = max(Sepal.Width),
            Sepal.Length_mean = mean(Sepal.Length),
            Sepal.Length_max = max(Sepal.Length))
```

`ungroup()`
===

To remove groupings structure from a data frame or a tibble.

```{r}
iris %>%
  group_by(Species) %>%
  ungroup() %>%
  summarize(Sepal.Width = mean(Sepal.Width),
            Petal.Width = mean(Petal.Width))
```



Join operations
===

A "join" operation combines two data sets. There are 4 types of join operations.

- **Inner join** (or just **join**): keeps just the rows each table that match the condition.
- **Left outer join** (or just **left join**): keeps all rows in the first table, and just the rows in the second table that match the condition.
- **Right outer join** (or just **right join**): keeps just the rows in the first table that match the condition, and all rows in the second table.
- **Full outer join** (or just **full join**): keeps all rows in both tables.

Note Column values that cannot be filled in are assigned `NA` values.



Illustration with two simple data sets.
===

```{r}
tab1_age <- data.frame(name = c("Ann", "Jenny", "Andrew"), 
                  age = c(70, 52, 40),
                  stringsAsFactors = FALSE)
tab2_testresult <- data.frame(name = c("Ann", "Nick", "Anderw"),
                  result = c("negative", "positive", "negative"),
                  stringsAsFactors = FALSE)
tab1_age
tab2_testresult
```

`inner_join()`
===

name column is common to both `tab1_age` and `tab2_testresult`. This keeps only the common rows (intersection) in both datasets.

```{r}
inner_join(x = tab1_age, y = tab2_testresult, by = "name")
```




`left_join()`
===

This keeps all names from `tab1_age`.

```{r}
left_join(x = tab1_age, y = tab2_testresult, by = c("name" = "name"))
```

`right_join()`
===

This keeps all names from `tab2_testresult`.

```{r}
right_join(x = tab1_age, y = tab2_testresult, by = "name")
```

`full_join()`
===

This keeps all rows from both data frames.

```{r}
full_join(x = tab1_age, y = tab2_testresult, by = "name")
```


Summary
===

- `tidyr` is a package for manipulating the structure of data frames
- `pivot_longer()`: make wide data longer
- `pivot_wider()`: make long data wider
- `unite()` and `separate()`: combine or split columns
- `dplyr` has advanced functionality that mirrors SQL
- `group_by()`: create groups of rows according to a condition
- `summarize()`: apply computations across groups of rows
- `*_join()` where `*` = `inner`, `left`, `right`, or `full`: join two data frames together according to common values in certain columns, and `*` indicates how many rows to keep.


