---
title: "Gapminder"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}
library(flexdashboard)
```
 


```{r message=FALSE, warning=FALSE, comment=NA}
library(gapminder)
library(tidyverse)
library(magrittr)
library(plotly)
library(gganimate)
library(ggplot2)
```


#Overview

column
------------------------------------------------------------------------
###Number of countries


```{r}
categories <- unique(gapminder$country) 
numberOfCountries <- length(categories)
valueBox(numberOfCountries, icon = "fa-pencil")
```


###Number of Continenets

```{r}
categories <- unique(gapminder$continent) 
numberOfCategories <- length(categories)
valueBox(numberOfCategories, icon = "fa-pencil", color = "red")
```




### Gapminder data

An excerpt of the data available at  [Gapminder](https://www.gapminder.org/data/). For each of 142 countries, the package provides values for life expectancy, GDP per capita, and population, every five years, from 1952 to 2007. \
\
Source : Jennifer Bryan (NA). gapminder: Data from Gapminder \
         https://github.com/jennybc/gapminder \
         http://www.gapminder.org/data/ \
         https://doi.org/10.5281/zenodo.594018 \
\
\
\
\
\
Created by W. Thashikala .D.Mendis \
(Done as a part of [STA 479 2.0 Special Topics in Statistics](https://tstdataviz.netlify.app/schedule/))

column
-----------------------------------------------------

###Composition of continent
```{r}
gapminder %>% 
  plot_ly(x = ~continent , color = gapminder$continent) %>% 
  layout(title="Composition of Continent")


```





#World Map

```{r}
library(maptools)
library(tibble)
library(ggrepel)
library(png)
library(grid)
library(sp)
library(dplyr)
library(tibble)
library(coronavirus)
library(ggthemes)
library(scales)
library(geofacet)
data("wrld_simpl")

coronaNew1 <- select(coronavirus, lat, long, country)
gapNew <- select(gapminder, pop,country, continent)
new <- inner_join(coronaNew1, gapNew, "country")

p <- ggplot() +
    geom_polygon(
      data = wrld_simpl,
      aes(x = long, y = lat, group = group), fill = "light blue", colour = "white")

 p +
   geom_point(aes(new$long, new$lat, size = pop, color = new$continent), data = new) +        scale_size_area()  + ggthemes::theme_map() + ggtitle("population")
 
 
```




#Distributions

column{.tabset .tabset-fade}
-----------------------------------------------------------------------
###Distributions of Life Expectancy by Continent

```{r}
p1 <- plot_ly(gapminder, x = gapminder$continent, y = gapminder$lifeExp)

p2 <- p1 %>% add_boxplot(color=~continent)
p2
```

###Distributions of gdpPercap by Continent

```{r}
p1 <- plot_ly(gapminder, x = gapminder$continent, y = gapminder$gdpPercap)

p2 <- p1 %>% add_boxplot(color=~continent)
p2
```


###Distributions of population by Continent

```{r}
p1 <- plot_ly(gapminder, x = gapminder$continent, y = gapminder$pop)

p2 <- p1 %>% add_boxplot(color=~continent)
p2
```


column{.tabset .tabset-fade}
------------------------------------------------

###Distributions of population by Life Expectancy and continent

```{r}

gapminderNEw <- gapminder %>% 
  mutate(Life_Expectancy = ifelse(lifeExp > 50, "High", "Low"))

p <- gapminderNEw %>%
    plot_ly(y = ~pop, color = I("black"), alpha = 0.1) 
p1 <- p %>% add_boxplot(y = ~pop, 
                          x = ~interaction(gapminderNEw$Life_Expectancy, continent),
                          color=~continent) 
p1

```



###Distributions of gdpPercap by Life Expectancy and continent

```{r}

gapminderNEw <- gapminder %>% 
  mutate(Life_Expectancy = ifelse(lifeExp > 50, "High", "Low"))

p <- gapminderNEw %>%
    plot_ly(y = ~gdpPercap, color = I("black"), alpha = 0.1) 
p1 <- p %>% add_boxplot(y = ~gdpPercap, 
                          x = ~interaction(gapminderNEw$Life_Expectancy, continent),
                          color=~continent)
p1

```

#Relations

column{data-width=650}
-------------------------------------------------------------------------


####Relationship between gdpPercap and life Expectancy

```{r}

p <- ggplot(gapminder, aes(gdpPercap, lifeExp, color = continent, frame = year)) +
  geom_point(aes(size = pop, ids = country)) + 
   geom_smooth(method = "lm", se=FALSE) +
  scale_x_log10() +
  theme_bw()

ggplotly(p)
```

column{.tabset .tabset-fade}
------------------------------------------------


####Relationship between gdpPercap and population 

```{r}

p <- ggplot(gapminder, aes(gdpPercap, pop, color = continent, frame = year)) +
  geom_point(aes(size = pop, ids = country)) + 
   geom_smooth(method = "lm", se=FALSE) +
  scale_x_log10() +
  theme_bw()

ggplotly(p)
```

column{.tabset .tabset-fade}
------------------------------------------------

####Relationship between gdpPercap and life Expectancy


```{r}

p <- ggplot(gapminder, aes(pop, lifeExp, color = continent, frame = year)) +
  geom_point(aes(size = pop, ids = country)) + 
   geom_smooth(method = "lm", se=FALSE) +
  scale_x_log10() +
  theme_bw()

ggplotly(p)
```


